---
title: "Data Science Homework 1"
author: "ak5357"
date: "2024-09-19"
output: github_document
---

```{r setup, echo=FALSE, message=FALSE}
library(tidyverse)
library(ggplot2)
```


## Problem 1

**Loading the penguins dataset and getting a `glimpse` of it.**
```{r}
data("penguins", package = "palmerpenguins")
glimpse(penguins)
```

**Describing the Data**

The `penguins` dataset has `r nrow(penguins)` rows and `r ncol(penguins)` columns, and contains information about the penguin characteristics listed below.

```{r get_range_str, echo=FALSE, message=FALSE}
get_range_str = function(x_list){
  output = paste("_Range:_ ", min(x_list, na.rm = TRUE), " to ", max(x_list, na.rm = TRUE))
  return(output)
}
```

| Variable | Unit | Data Type | Values |
|-----|-----|-----|-----|
| Species | NA |  Factor | `r levels(pull(penguins, species))` |
| Island | NA |  Factor | `r levels(pull(penguins, island))` |
| Bill Length | mm |  Double |  `r get_range_str(pull(penguins, bill_length_mm))` |
| Bill Depth | mm |  Double | `r get_range_str(pull(penguins, bill_depth_mm))` |
| Flipper Length | mm |  Integer | `r get_range_str(pull(penguins, flipper_length_mm))` |
| Body Mass | g |  Integer | `r get_range_str(pull(penguins, body_mass_g))` |
| Sex | NA |  Factor | `r levels(pull(penguins, sex))` |
| Year | NA |  Integer | `r get_range_str(pull(penguins, year))` |


**Calculating Mean Flipper Length**

```{r mean_flipper_length}
mean_flipper_length = 
  pull(penguins, flipper_length_mm) |> 
  mean(na.rm = TRUE) |> 
  round(2)

```

The mean flipper length among the penguins in the dataset is `r mean_flipper_length` mm.

**Scatterplot of Flipper Length vs Bill Length**

```{r scatterplot, warning = FALSE}
ggplot(penguins, aes(x = flipper_length_mm, y = bill_length_mm, color = species)) + 
  geom_point() + 
  labs(title = "Flipper Length (mm) vs. Bill Length (mm)",
       x = "Bill Length (mm)",
       y = "Flipper Length (mm)",
       color = "Species"
       )
```


**Save Scatterplot to Project Directory**
```{r save_plot, warning = FALSE}
ggsave(filename = "penguins_scatterplot.png")
```

# Problem 2

**Create a dataframe**
```{r sample_dataframe}
df = tibble(
  norm_sample = rnorm(10),
  greater_than_0 = norm_sample > 0,
  char = stringi::stri_rand_strings(10, 1, pattern = "[a-z]"),
  fac = factor(sample(c("bird", "cat", "dog"), size = 10, replace = TRUE))
)

df
```

**Taking the mean of each variable**
```{r}
mean(pull(df, norm_sample))

mean(pull(df, greater_than_0))

mean(pull(df, char))

mean(pull(df, fac))
```

Taking the mean of my variables results in numeric values for the first two columns, `norm_samp` and `greater_than_0`, and warning messages for the latter columns, `char` and `fac`.

This exercise and the warning messages above demonstrate that the `mean()` function can only be used on numeric or logical vectors. Applying the `mean()` function



For the `norm_samp` column, as expected the output value is the arithmetic mean of the numeric values in the column. For `greater_than_0`, the output value is a probabilistic mean, showing what proportion of the column's values are `TRUE`. Both `char` and `fac` have an output of `NA` because their arguments are not numeric or logical.












