---
title: "Data Science Homework 1"
author: "ak5357"
date: "2024-09-19"
output: github_document
---

```{r setup, echo=FALSE, message=FALSE}
library(tidyverse)
library(ggplot2)
```


## Problem 1

**Loading the penguins dataset and getting a `glimpse` of it.**
```{r}
data("penguins", package = "palmerpenguins")
glimpse(penguins)
```

**Describing the Data**

The `penguins` dataset has `r nrow(penguins)` rows and `r ncol(penguins)` columns, and contains information about the attributes listed below.

```{r get_range_str, echo=FALSE, message=FALSE}
get_range_str = function(x_list){
  output = paste(
    "_Range:_ ",
    min(x_list, na.rm = TRUE),
    " to ",
    max(x_list, na.rm = TRUE)
    )
  return(output)
}
```

| Variable | Column Name | Unit | Data Type | Values |
|-----|-----|-----|-----|-----|
| Species | species | N/A |  Factor | `r levels(pull(penguins, species))` |
| Island | island | N/A |  Factor | `r levels(pull(penguins, island))` |
| Bill Length | bill_length_mm | mm |  Double |  `r get_range_str(pull(penguins, bill_length_mm))` |
| Bill Depth | bill_depth_mm | mm |  Double | `r get_range_str(pull(penguins, bill_depth_mm))` |
| Flipper Length | flipper_length_mm | mm |  Integer | `r get_range_str(pull(penguins, flipper_length_mm))` |
| Body Mass | body_mass_g | g |  Integer | `r get_range_str(pull(penguins, body_mass_g))` |
| Sex | sex | N/A |  Factor | `r levels(pull(penguins, sex))` |
| Year | year | N/A |  Integer | `r get_range_str(pull(penguins, year))` |


**Calculating Mean Flipper Length**

```{r mean_flipper_length}
mean_flipper_length = 
  pull(penguins, flipper_length_mm) |> 
  mean(na.rm = TRUE) |> 
  round(2)

```

The mean flipper length among the penguins in the dataset is `r mean_flipper_length` mm.

**Scatterplot of Flipper Length vs Bill Length**

```{r scatterplot, warning = FALSE}
ggplot(penguins, aes(x = flipper_length_mm, y = bill_length_mm, color = species)) + 
  geom_point() + 
  labs(title = "Flipper Length (mm) vs. Bill Length (mm)",
       x = "Bill Length (mm)",
       y = "Flipper Length (mm)",
       color = "Species"
       )
```


**Save Scatterplot to Project Directory**
```{r save_plot, warning = FALSE}
ggsave(filename = "penguins_scatterplot.png")
```

# Problem 2

**Create a dataframe**
```{r sample_dataframe}
df = tibble(
  norm_sample = rnorm(10),
  greater_than_0 = norm_sample > 0,
  char = stringi::stri_rand_strings(10, 1, pattern = "[a-z]"),
  fac = factor(sample(c("bird", "cat", "dog"), size = 10, replace = TRUE))
)

df
```

**Taking the mean of each variable**
```{r mean_function}
mean(pull(df, norm_sample))

mean(pull(df, greater_than_0))

mean(pull(df, char))

mean(pull(df, fac))
```

This exercise and the warning messages above demonstrate that the `mean()` function can only be used on numeric or logical vectors. Applying the `mean()` function on the `norm_samp` or `greater_than_0` columns yields their arithmetic and probabilistic means, respectively. However, applying the `mean()` function on either `char` or `fac` yields an output of `NA`, since the values in those columns are neither numeric nor logical.


```{r as_numeric, eval = FALSE}
as.numeric(pull(df, norm_sample))

as.numeric(pull(df, greater_than_0))

as.numeric(pull(df, char))

as.numeric(pull(df, fac))
```

When I pass each column through the `as.numeric()` function, the output is as follow:

* `norm_sample`: This column does not appear to change at all, since the values are already numeric.

* `greater_than_0`: All `TRUE` values are converted to `1`, and all `FALSE` values are converted to `0`.

* `char`: Returns an array of `NA` values.

* `fac`: Each factor is converted to either `1`, `2`, or `3` based on their order when I defined them.

This explains what happens when we pass numeric, logical, or character arrays into the `mean()` function. However, in my opinion, it does not directly explain why we receive a `NA` value when passing an array of factors to the `mean()` function.
