---
title: "Data Science Homework 1"
author: "ak5357"
date: "2024-09-19"
output: github_document
---

```{r setup, echo=FALSE, message=FALSE}
library(tidyverse)
```


## Problem 1

Loading the penguins dataset and getting a `glimpse` of it.
```{r}
data("penguins", package = "palmerpenguins")
glimpse(penguins)
```

**Describing the Data**

The `penguins` dataset has `r nrow(penguins)` rows and `r ncol(penguins)` columns, and contains information about the penguin characteristics listed below.

```{r get_range_str, echo=FALSE, message=FALSE}
get_range_str = function(x_list){
  output = paste("_Range:_ ", min(x_list, na.rm = TRUE), " to ", max(x_list, na.rm = TRUE))
  return(output)
}
```

| Variable | Unit | Data Type | Values |
|-----|-----|-----|-----|
| Species | NA |  Factor | `r levels(pull(penguins, species))` |
| Island | NA |  Factor | `r levels(pull(penguins, island))` |
| Bill Length | mm |  Double |  `r get_range_str(pull(penguins, bill_length_mm))` |
| Bill Depth | mm |  Double | `r get_range_str(pull(penguins, bill_depth_mm))` |
| Flipper Length | mm |  Integer | `r get_range_str(pull(penguins, flipper_length_mm))` |
| Body Mass | g |  Integer | `r get_range_str(pull(penguins, body_mass_g))` |
| Sex | NA |  Factor | `r levels(pull(penguins, sex))` |
| Year | NA |  Integer | `r get_range_str(pull(penguins, year))` |


**Calculating Mean Flipper Length**

```{r}
mean_flipper_length = 
  pull(penguins, flipper_length_mm) |> 
  mean(na.rm = TRUE) |> 
  round(2)

```

The mean flipper length among the penguins in the dataset is `r mean_flipper_length` mm.

**Scatterplot of Flipper Length vs Bill Length**







